---
title: "MLM analysis for P2"
author: "Hannah"
date: "11/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
require(lme4)
require(lmerTest)
require(dplyr)
require(knitr)
require(ggplot2)
require(kableExtra)

dat = read.delim("AllSubs_acceptedTrials_long_nobe_nobs.txt")
```

*Redone with new set of 11 electrodes. Reblocked participants who were missing trials and increased number of rejected electrodes.*

## 1. Gender categorization task


``` {r}
gendat = filter(dat, Task == "Gender")

gendat_temp = gendat
gendat_temp$Condition = paste(gendat_temp$TarRace, gendat_temp$TarGender, gendat_temp$Fix, sep = "_")

numTrials = NULL
for (i in unique(gendat_temp$Subject)) {
  temp = filter(gendat_temp, Subject == i)
  for (c in unique(gendat_temp$Condition)) {
    num = length(unique(temp$Trial[temp$Condition == c]))
    numTrials = rbind(numTrials, data.frame(Subject = i, Condition = c, numTrials = num))
    }
}

```
Median number of trials per condition per participant: `r median(numTrials$numTrials)`  
Min:`r min(numTrials$numTrials)`  
Max:`r max(numTrials$numTrials)`  




### 1a. P2 waveforms

See figures in "4 Grand averages/Plots".


### 1b. Effects of Target Race, Target Gender, Fixation, and Participant Race on P2 amplitude (all trials)
DV is trial level data. Subjects with < 50% trials accepted are excluded.  
Electrodes: `r unique(gendat$Electrode)`  

P2 is quantified as mean amplitude from 130-190 ms.  

Participant race, target race, and target gender are effect coded. Both eye and forehead fixation trials are included.  

Included n = 63 (32 black, 31 white)  

```{r}
# 2. Look at Gender task with Par Race, Target Race, and Target Gender as predictors

# Determine model (commented out to make knitting faster)
# n1 = lmer(value ~ 1 + (TarRace*TarGender|Subject) + (1|Electrode), data = gendat)
# n2 = lmer(value ~ 1 + (TarRace+TarGender|Subject) + (1|Electrode), data = gendat)
# n3 = lmer(value ~ 1 + (TarRace|Subject) + (1|Electrode), data = gendat) # use this

# Add predictors, effect coding
gendat$TarRace.e = -1
gendat$TarRace.e[gendat$TarRace == "White"] = 1

gendat$TarGender.e = -1
gendat$TarGender.e[gendat$TarGender == "male"] = 1

gendat$ParRace.e = -1
gendat$ParRace.e[gendat$ParRace == "White"] = 1

gendat$Fix.e = -1
gendat$Fix.e[gendat$Fix == "forehead"] = 1

n4 = lmer(value ~ TarRace.e*TarGender.e*Fix.e*ParRace.e + (TarRace.e + TarGender.e|Subject) + (1|Electrode), data = gendat)
sum.n4 = summary(n4)
```

``` {r}
sum.n4$call
sum.n4$varcor
kable(round(sum.n4$coefficients, digits=3)) %>% 
  kable_styling(full_width = F, position = "center", bootstrap_options = "condensed") %>% 
  row_spec(c(1:5,7,11:12,14), bold = T, background = "goldenrod2")
```

### 1c. Just Black participants  
Fixation and target race are effect coded.
``` {r}
n6 = lmer(value ~ TarRace.e*Fix.e + (TarRace.e + Fix.e|Subject) + (1|Electrode), data = filter(gendat, ParRace == "Black"))
sum.n6 = summary(n6)
sum.n6$call
sum.n6$varcor
kable(round(sum.n6$coefficients, digits=3)) %>% 
  kable_styling(full_width = F, position = "center", bootstrap_options = "condensed") %>% 
  row_spec(c(1,3:4), bold = T, background = "goldenrod2")
```

### 1d. Just White participants  
Fixation and target race are effect coded. 
``` {r}
n6 = lmer(value ~ TarRace.e*Fix.e + (TarRace.e + Fix.e|Subject) + (1|Electrode), data = filter(gendat, ParRace == "White"))
sum.n6 = summary(n6)
sum.n6$call
sum.n6$varcor
kable(round(sum.n6$coefficients, digits=3)) %>% 
  kable_styling(full_width = F, position = "center", bootstrap_options = "condensed") %>% 
  row_spec(c(1,4), bold = T, background = "goldenrod2")
```

### 1e. Plot for interactions

``` {r}
facet_labels = c(Black = "Black participants", White = "White participants")
ggplot(gendat, aes(Fix, value, fill = TarRace)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", position = position_dodge(width=.9), width = .2) +
  facet_wrap(~ParRace, labeller=labeller(ParRace = facet_labels)) + 
  #  ggtitle("Total number of errors") +
  labs(y = "P2 amplitude") +
  scale_fill_manual(values=c("black","grey70"), guide = guide_legend(title = "Target Race")) +
  theme_bw() +
  ggtitle("Gender task") +
  #coord_cartesian(ylim = c(400, 500)) +
  theme(panel.grid.major = element_line(color = "white"),
        panel.grid.minor = element_line(color = "white"),
        strip.text.x = element_text(face = "bold", size = 14),
        strip.background = element_rect(fill = "grey98"),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        plot.title = element_text(hjust = 0.5))
```


## 2. Race categorization task  

``` {r}
racedat = filter(dat, Task == "Race")

racedat_temp = racedat
racedat_temp$Condition = paste(racedat_temp$TarRace, racedat_temp$TarGender, racedat_temp$Fix, sep = "_")

numTrials = NULL
for (i in unique(racedat_temp$Subject)) {
  temp = filter(racedat_temp, Subject == i)
  for (c in unique(racedat_temp$Condition)) {
    num = length(unique(temp$Trial[temp$Condition == c]))
    numTrials = rbind(numTrials, data.frame(Subject = i, Condition = c, numTrials = num))
    }
}


```
Median number of trials per condition per participant: `r median(numTrials$numTrials)`  
Min:`r min(numTrials$numTrials)`  
Max:`r max(numTrials$numTrials)`  

### 2a. P2 waveforms

See figures in "4 Grand averages/Plots".

### 2b. Effects of Target Race, Target Gender, and Participant Race on P2 amplitude (all trials)

DV is trial level data. Subjects with < 50% trials accepted are excluded.  
Electrodes: `r unique(racedat$Electrode)` 

P2 is quantified as mean amplitude from 130-190 ms.  

Participant race, target race, and target gender are effect coded. Both eye and forehead fixation trials are included.  

Included n = 63 (33 black, 30 white)


```{r}
# 1. Look at Race Task with Par Race, Target Race, and Target Gender as predictors

# Determine model (commented out to make kitting faster)
# m1 = lmer(value ~ 1 + (TarRace*TarGender|Subject) + (1|Electrode), data = racedat)
# m2 = lmer(value ~ 1 + (TarRace+TarGender|Subject) + (1|Electrode), data = racedat) # use this

# Add predictors, effect coding
racedat$TarRace.e = -1
racedat$TarRace.e[racedat$TarRace == "White"] = 1

racedat$TarGender.e = -1
racedat$TarGender.e[racedat$TarGender == "male"] = 1

racedat$ParRace.e = -1
racedat$ParRace.e[racedat$ParRace == "White"] = 1

racedat$Fix.e = -1
racedat$Fix.e[racedat$Fix == "forehead"] = 1

m3 = lmer(value ~ TarRace.e*TarGender.e*Fix.e*ParRace.e + (TarRace.e+TarGender.e|Subject) + (1|Electrode), data = racedat) # use this
sum.m3 = summary(m3)
```
All variables are effect coded.  


``` {r}
sum.m3$call
sum.m3$varcor
kable(round(sum.m3$coefficients, digits=3)) %>% 
  kable_styling(full_width = F, position = "center", bootstrap_options = "condensed") %>% 
  row_spec(c(1:2,4:7, 12, 14), bold = T, background = "goldenrod2")
```


### 2c. Just Black participants  
Fixation and target race are effect coded. 
``` {r}
m5 = lmer(value ~ TarRace.e*Fix.e + (TarRace.e + Fix.e|Subject) + (1|Electrode), data = filter(racedat, ParRace == "Black"))
sum.m5 = summary(m5)
sum.m5$call
sum.m5$varcor
kable(round(sum.m5$coefficients, digits=3)) %>% 
  kable_styling(full_width = F, position = "center", bootstrap_options = "condensed") %>% 
  row_spec(c(1:4), bold = T, background = "goldenrod2")
```

### 2d. Just White participants 
Fixation and target race are effect coded. 
``` {r}
m5 = lmer(value ~ TarRace.e*Fix.e + (TarRace.e + Fix.e|Subject) + (1|Electrode), data = filter(racedat, ParRace == "White"))
sum.m5 = summary(m5)
sum.m5$call
sum.m5$varcor
kable(round(sum.m5$coefficients, digits=3)) %>% 
  kable_styling(full_width = F, position = "center", bootstrap_options = "condensed") %>% 
  row_spec(c(1:4), bold = T, background = "goldenrod2")
```

### 1e. Plot for interactions

``` {r}
facet_labels = c(Black = "Black participants", White = "White participants")
ggplot(racedat, aes(Fix, value, fill = TarRace)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", position = position_dodge(width=.9), width = .2) +
  facet_wrap(~ParRace, labeller=labeller(ParRace = facet_labels)) + 
  #  ggtitle("Total number of errors") +
  labs(y = "P2 amplitude") +
  scale_fill_manual(values=c("black","grey70"), guide = guide_legend(title = "Target Race")) +
  theme_bw() +
  ggtitle("Race task") +
  #coord_cartesian(ylim = c(400, 500)) +
  theme(panel.grid.major = element_line(color = "white"),
        panel.grid.minor = element_line(color = "white"),
        strip.text.x = element_text(face = "bold", size = 14),
        strip.background = element_rect(fill = "grey98"),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        plot.title = element_text(hjust = 0.5))
```


## 3. Comparison of tasks

DV is trial level data. Subjects with < 50% trials accepted are excluded. 

Included n = 65 (34 black, 31 white)

#### Testing difference in effect of RACE and GENDER across tasks, collapsing across participant race:  
Target Race, Target Gender, and Task are included as effect-coded predictors. Only the Target Race x Task and Target Gender x Task interactions are included. Both eye and forehead fixation trials are included.  
```{r}

# Add predictors, effect coding
dat$TarRace.e = -1
dat$TarRace.e[dat$TarRace == "White"] = 1

dat$Task.e = -1
dat$Task.e[dat$Task == "Race"] = 1

dat$TarGender.e = -1
dat$TarGender.e[dat$TarGender == "male"] = 1

m3 = lmer(value ~ TarGender.e*Task.e + TarRace.e*Task.e + (TarGender.e+TarRace.e + Task.e|Subject) + (1|Electrode), data = dat) # use this
sum.m3 = summary(m3)

sum.m3$call
sum.m3$varcor
kable(round(sum.m3$coefficients, digits=3))%>% 
  kable_styling(full_width = F, position = "center", bootstrap_options = "condensed") %>% 
  row_spec(c(1:2, 4:5), bold = T, background = "goldenrod2")
```
